---
fontsize: 12pt
geometry: margin=0.50in
subparagraph: yes
params: 
    title:
      value: x
    date:
      value: x
    sampleDir:
      value: x
    samplename:
      value: x

title: "`r params$title`"
date:  "`r params$date`"
sampleDir: "`r params$sampleDir`"
samplename: "`r params$samplename`"
header-includes:
   - \usepackage{pdflscape}
   - \newcommand{\blandscape}{\begin{landscape}}
   - \newcommand{\elandscape}{\end{landscape}}
   - \usepackage{pdfpages}
   - \usepackage{booktabs}
   - \usepackage{longtable}
   - \usepackage{makecell}
   - \usepackage{float}
output:
  pdf_document:
    toc: false
    keep_tex: no
    latex_engine: xelatex
---

\vspace{0.5cm}

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4}
library(ggplot2)
library(knitr)
library(kableExtra)
library(stringr)
library(dplyr)
library(tidyverse)
library(scales)

#https://www.dataquest.io/blog/r-markdown-tips-tricks-and-shortcuts/#:~:text=To%20break%20a%20line%20in,spaces%20and%20then%20hit%20return%20.

sampleDir <- params$sampleDir
samplename <- params$samplename

##test inputs start
#sampleDir <- '/home/agmcfarland/flu_project/FluPipeline/run_test/output/sampleOutputs/H1N1_A_Brisbane_59_2007_snp'
#samplename <- 'H1N1_A_Brisbane_59_2007_snp'
#library(optparse)
#setwd('/home/agmcfarland/flu_project/FluPipeline/run_test/output/sampleOutputs/H1N1_A_Brisbane_59_2007_snp')
##test inputs end

# load in Rdata object
if (file.exists(paste0(sampleDir,'/',samplename,'.Rdata'))){
  load(paste0(sampleDir,'/',samplename,'.Rdata'))
} else {
  print('no Rdata')
}


# helper functions
get_VarCounts <- function(tablename){
  # get counts of variants per segment for a variantTable
  # return a datafarame of segment id, segment, count, and arrange by segment number.
  df <- read.csv(paste0(sampleDir,'/',tablename))%>%
    dplyr::filter(!ALT%in%c('e','?'))%>%
    dplyr::group_by(CHROM)%>%
    dplyr::mutate(count_vars=length(POS))%>%
    dplyr::ungroup()%>%
    dplyr::mutate(segment=stringr::str_sub(CHROM,start=-1))%>%
    dplyr::rename(segment_id=CHROM)%>%
    dplyr::select(segment_id,segment,count_vars)%>%
    unique()%>%
    dplyr::arrange(segment)
}

# get variant counts

df_allvar <- tryCatch(
    expr = {
      get_VarCounts(tablename='variantTable.csv')
    },
    error = function(e){ 
      return(data.frame('no variants'=NA))
    }
)

df_majorvar <- tryCatch(
    expr = {
      get_VarCounts(tablename='variantTableMajor.csv')
    },
    error = function(e){ 
      return(data.frame('no Major variants'=NA))
    }
)

# get coverage calculations
df_cov <- merge(opt$refGenomePercentCovered,opt$refGenomePercentCovered_5reads,by='V1',all.x=TRUE,all.y=TRUE)
df_cov <- df_cov%>%
  dplyr::mutate(segment=stringr::str_sub(V1,start=-1))%>%
  dplyr::rename(segment_id=V1)%>%
  dplyr::select(segment_id,segment,perc_gt_1,perc_gt_5)%>%
  dplyr::mutate(perc_gt_1 = round(perc_gt_1,2), perc_gt_5 = round(perc_gt_5,2))%>%
  dplyr::arrange(segment)

# get reference genome
reference_genome <- opt$refGenomeFasta

# reference strain coverage summary
df_ref <- read.csv(paste0(sampleDir,'/','reference_coverage_',samplename,'.csv'))%>%
  select(-sample)%>%
  mutate(segment_coverage_average=round(segment_coverage_average,2),
         segment_average_read_depth=round(segment_average_read_depth,2),
         reference=base::sub('\\.fasta.*', '', reference)) #capture everything before .fasta

# per-segment coverage for each reference
df_refstrains <- read.csv(paste0(sampleDir,'/','combined_coverage_stats_',samplename,'.csv'))

df_refstrains <- df_refstrains%>%
  mutate(average_coverage=round(segment_coverage,2),
         average_depth=round(average_read_depth,2))%>%
  select(-segment_coverage,-average_read_depth)%>%
  pivot_wider(names_from=reference, values_from=c(average_coverage,average_depth))


## get per-base depth values from pileup output file
df_pile <- tryCatch({
  opt$pileupData
}, error = function(e) {
  return(data.frame())
})

df_pile <- df_pile%>%select(1,2,3,4,5)%>%
  dplyr::mutate(segment=stringr::str_sub(V1,start=-1))

# make a per segment depth summary table. To merge with df_cov
df_pilestats <- df_pile%>%
  dplyr::group_by(segment)%>%
  dplyr::mutate(mean_depth = round(mean(V4),2),
         median_depth = round(median(V4),2),
         min_depth = min(V4),
         max_depth = max(V4),
         std_depth = round(sd(V4),2))%>%
  dplyr::select(segment,V1,mean_depth,median_depth,min_depth,max_depth,std_depth)%>%
  ungroup()%>%
  unique()%>%
  dplyr::arrange(segment)%>%
  dplyr::select(-segment)

## get tsv
if ( file.exists(paste0(sampleDir,'/','nextclade_output','/',samplename,'_clade_assignment.tsv')) ){
  df_nc <- read.csv(paste0(sampleDir,'/','nextclade_output','/',samplename,'_clade_assignment.tsv'), sep='\t')%>%
    select(1:13)
}else{
  df_nc <- data.frame('No nextclade assignment available'=NA)
}



```

```{r, echo=FALSE, results='asis'}
# Prints the reference genome that is used 
cat(paste('**Reference Genome: **',basename(opt$refGenomeFasta),'  \n'))

if ( ncol(df_nc) == 13 ){
  cat(paste('**Clade assignment: **', df_nc$clade))
} else {
  cat(paste('**Clade assignment: **', 'No nextclade assignment available'))
}

```

# Coverage and Depth
\small
Table 1: Percentage of genome covered by 1 or 5 reads (perc_gt_1, perc_gt_5) and and per-base read depth statistics per segment.
\normalsize
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=2}

df_cov1 <- merge(df_cov, df_pilestats, by.x='segment_id',by.y='V1',all.x=TRUE)%>%
  arrange(segment)

kable(df_cov1, 'latex', booktabs = T, linesep = '', align = 'c') %>% 
  kable_styling(font_size = 7.5, latex_options="hold_position")
```

# Variants
\small
Table 2: Counts of all variants per segment.
\normalsize
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=2}

kable(df_allvar, 'latex', booktabs = T, linesep = '', align = 'c') %>% 
  kable_styling(font_size = 7.5, latex_options="hold_position")

```

\small
Table 4: Counts of major variants per segment (>=50% of reads contain variant).
\normalsize
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=2}

kable(df_majorvar, 'latex', booktabs = T, linesep = '', align = 'c') %>% 
  kable_styling(font_size = 7.5, latex_options="hold_position")

```
\newpage
\small
Figure 1: Read depth and coverage map.
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=10, fig.align='center'}

## Make a combined pileup and variant table
df_variants <- tryCatch(
    expr = {
      read.csv(paste0(sampleDir,'/variantTableMajor.csv'))%>%
        dplyr::select(CHROM,POS)%>%unique()%>%
        dplyr::rename('segment_name'='CHROM',
                  'var_position'='POS')%>%
        dplyr::mutate(segment_name=as.character(segment_name),
                      var_position=as.numeric(var_position),
                      var_position2=var_position)
    },
    error = function(e){ 
      # return a dataframe with dummy data so that plottng won't be affected.
      return(data.frame('segment_name'=as.character('none'),'var_position'=as.numeric(0), 'var_position2'=as.numeric(0)))
    }
)

df_pile1 <- df_pile%>%
  dplyr::rename('segment_name'=1,'position'=2,'original'=3,'depth'=4,'quality'=5)%>%
  dplyr::select(-quality,-original)%>%
  dplyr::mutate(segment_name=as.character(segment_name),
                position=as.numeric(position),
                depth=as.numeric(depth),
                segment=as.numeric(segment))

# merge pileup positions with variant positions
df_pile1 <- merge(df_pile1, df_variants, by.x=c('segment_name','position'), by.y=c('segment_name','var_position'), all.x=TRUE)

ggplot(df_pile1, aes(x=position, y=depth, group=segment_name))+
  facet_wrap(~segment, scales='free_x',ncol=1, strip.position='right')+
  geom_area()+
  #geom_bar(stat='identity')+
  geom_point(aes(x=var_position2, y=3,group=segment), color='yellow',alpha=0.85)+
  theme_bw()+
 theme(panel.grid.major = element_blank(),
       panel.grid.minor = element_blank(),
       panel.background = element_blank(),
       panel.spacing = unit(0, 'lines'),
       strip.text = element_text(size=14, colour='gray50'),
       strip.text.y = element_text(angle=0),
       strip.background = element_blank(), # getting rid of the panel to declutter
       axis.line = element_line(colour = 'black'),
       axis.text.y = element_text(size=10),
       axis.text.x = element_text(size=7),
       legend.position='none',
       legend.title = element_blank())+
  scale_x_continuous(labels = function(x)x/1000, # converting bp tick mark number to kilobasepairs
                     breaks=seq(from=0,to=3000,by=250),
                     expand=c(0,0))+
  ylab(label = expression(Read~Coverage))+
  xlab(label = expression(Genomic~Position~(Kb)))+
  geom_hline(aes(yintercept=median(df_pile1$depth)), linetype='dashed',colour='lightblue')+
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x)))

```

\newpage
\small
Figure 2: Read depth and coverage map zoomed at a depth of 100 reads.
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=10, fig.align='center'}

ggplot(df_pile1, aes(x=position, y=depth, group=segment_name))+
  facet_wrap(~segment, scales='free_x',ncol=1, strip.position='right')+
  #geom_bar(stat='identity')+
  geom_area()+
  geom_point(aes(x=var_position2, y=3,group=segment), color='yellow',alpha=0.85)+
  theme_bw()+
 theme(panel.grid.major = element_blank(),
       panel.grid.minor = element_blank(),
       panel.background = element_blank(),
       panel.spacing = unit(0, 'lines'),
       strip.text = element_text(size=14, colour='gray50'),
       strip.text.y = element_text(angle=0),
       strip.background = element_blank(), # getting rid of the panel to declutter
       axis.line = element_line(colour = 'black'),
       axis.text.y = element_text(size=10),
       axis.text.x = element_text(size=7),
       legend.position='none',
       legend.title = element_blank())+
  scale_x_continuous(labels = function(x)x/1000, # converting bp tick mark number to kilobasepairs
                     breaks=seq(from=0,to=3000,by=250),
                     expand=c(0,0))+
  ylab(label = expression(Read~Coverage))+
  xlab(label = expression(Genomic~Position~(Kb)))+
  geom_hline(aes(yintercept=median(df_pile1$depth)), linetype='dashed',colour='lightblue')+
  coord_cartesian(ylim=c(0, 100))+
  scale_y_continuous(breaks=c(0,20,40,60,80,100))

```


\newpage
# Clade assignment and HA sequence stats
\small
Table 4: A. Clade assignment given to the strain using Nextclade. Highlighting the quality (qc) scores and qualitative assignment. B. Number of SNPs/indels. C. Amino acid level effects of SNPs/indels.

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=2}
if ( ncol(df_nc)>1 ){
  kable(df_nc%>%
          select(1:4)
          , 'latex', booktabs = T, linesep = '', align = 'c') %>% 
    kable_styling(font_size = 7.5,latex_options="hold_position")
  
  kable(df_nc%>%
          select(5:8)
          , 'latex', booktabs = T, linesep = '', align = 'c') %>% 
    kable_styling(font_size = 7.5,latex_options="hold_position")
  
  kable(df_nc%>%
          select(9:11)
          , 'latex', booktabs = T, linesep = '', align = 'c') %>% 
    kable_styling(font_size = 7.5,latex_options="hold_position")
} else {
      kable(df_nc, 'latex', booktabs = T, linesep = '', align = 'c') %>% 
    kable_styling(font_size = 7.5,latex_options="hold_position")
  }
```

\newpage
# Reference strain selection
\small
Table 5: Summary of read coverage and depth for reference strain alignment using a randomized subset of reads.

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=2}

kable(df_ref, 'latex', booktabs = T, linesep = '', align = 'c') %>% 
  kable_styling(font_size = 7.5,latex_options="hold_position")

```
\small
Table 6: Per-segment read coverage for reference strain alignment using a randomized subset of reads.

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=2}

kable(df_refstrains%>%
       dplyr:: select(starts_with(c('segment','average_coverage')))%>%
       dplyr::rename_with(~str_remove(., 'average_coverage_'))
        , 'latex', booktabs = T, linesep = '', align = 'c') %>% 
  kable_styling(font_size = 7.5, latex_options="hold_position")

```

Table 7: Per-segment read depth for reference strain alignment using a randomized subset of reads.

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=2}

kable(df_refstrains%>%
       dplyr:: select(starts_with(c('segment','average_depth')))%>%
       dplyr::rename_with(~str_remove(., 'average_depth_'))
        , 'latex', booktabs = T, linesep = '', align = 'c') %>% 
  kable_styling(font_size = 7.5, latex_options="hold_position")

```
\newpage

# Software environment

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=2}
kable(opt$softwareVersionTable, 'latex',longtable = T, booktabs = T,  linesep = '', align = 'l') %>% 
  column_spec(1, width = "5cm") %>% 
  kable_styling(latex_options =c("repeat_header"))
```

